% !TeX spellcheck = en_EN-English

\section{Future record prediction}
\label{record_prediction}

%This task is both the most important and the most challenging to train. Our goal is to develop a model capable of predicting a patient's potential next record based on their previous ones. Generally, this task is nearly impossible with the amount of information available, as numerous factors influence whether, when, and what new disease a patient might contract, how their current state will evolve, and what specific actions a doctor will take.
%\\

This task is both the most important and the most challenging to train. Our goal is to develop a model capable of predicting a patient’s potential next record based on their previous ones. In general, this is nearly impossible given the available information, as many other factors influence if, when, and what new disease a patient might contract, how their condition will evolve, and what actions a doctor will take.
\\

Fortunately, our ultimate goal is not to predict a patient's specific future but to estimate the likely total cost of their future records. We anticipate that even if we cannot predict exact outcomes, we can still estimate the overall cost.
\\

To achieve this prediction, we experimented with multiple models. The first three models we tested were Recurrent Neural Networks (RNNs): multi-layer Elman RNN, multi-layer Gated Recurrent Unit (GRU) RNN, and multi-layer Long Short-Term Memory (LSTM) RNN. The architecture of these models is detailed in Sec. \ref{theoryRNN}. The last model we tried was a Transformer, specifically a Decoder-only Transformer model, explained in Sec. \ref{theoryTrans}.

