% !TeX spellcheck = en_EN-English

\chapter{Proposed Methods}
\label{propMet}

The task of predicting the future cost of a patient can be split into multiple sub-tasks, which follow each other. The sub-tasks are:

\begin{enumerate}
	\item Embed the patient’s history into numerical vectors
	\item Compute the expected number of records the patient will have in the next year
	\item Predict future records for the patient
	\item Predict the cost of each future record
	\item Compute the total cost of the patient for the next year
\end{enumerate}


\input{proposed methods/embedding}

\section{Prediction of future number of records}

Our task is to determine how much a patient will cost in the future, specifically in the next year. Since our approach predicts future records, assigns costs to them, and sums these into a total cost, we need to estimate how many records to generate to simulate the next year. For this task, we tried two different approaches.
\\

The first approach was to predict the approximate number of records using linear regression, which takes the counts of records from previous years and predicts the next year’s count. The second approach used a stopping criterion based on timestamp information available in each input and generated record. This method stops generating records once the difference between the last timestamp from the patient’s data and the last generated timestamp surpasses a one-year threshold.
\\

Each method has its own disadvantages. In the case of linear regression, the number of records per year can vary significantly. Even if we expect an increase \cite{num_of_vis}, regression might fail to capture this trend, especially if the patient’s data includes only a few years of historical records. A potential issue with the second approach is its reliance on how well the model learned that timestamps should always increase.

\input{proposed methods/future_records_prediction}

\input{proposed methods/cost_prediction}
\newpage

\section{Prediction of future cost of patient}
\label{pat_fut_pred_met}

This last task is our main goal: to predict the cost of a patient in the next year. To achieve this, we utilize the results from all our previous tasks. First, we embed the patient’s records. Then, we predict the records that the patient could receive in the next year. For each predicted record, we predict its cost category. Finally, all costs are summed and transformed into a patient cost category.
